Augmentation of a Historical Harpsichord Keyboard Replica for Haptic-Enabled Interaction in Museum Exhibitions

### 2025-02-04 src/article.tex
Start catalina comments

The \anon{Tagliavini} Collection in \anon{Bologna \cite{Tagliavini2007}}, renowned for its historical keyboard instruments, exemplifies this dilemma. With more than seventy early keyboard instruments, including many unique Renaissance Italian harpsichords, spinets and virginals, the collection stands out as a valuable resource for musicologists, organologists and musicians alike. Preserving the instruments' authenticity was the cornerstone of \anon{Ferdinando Tagliavini}’s vision. This guiding principle led him to collect only playable instruments, or instruments which could be restored to playing condition after minimal intervention.

However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—typically during concerts or by experienced performers with special permission from the curator. 
To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian Renaissance, which is the subject of this paper.
The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages. The interface is currently connected to a commercial software sampler, though the ultimate goal is to use sample instruments from the collection. Visitors to the museum are invited to play the interface and listen through a pair of headphones. This work outlines the technological aspects of the interface's construction and offers reflections on its role within the \anon{Tagliavini} Collection and potential application within the wider musical instrument museum context. 
% The paper outlines the design and motivations behind this replica keyboard interface.

@@

McAlpine discusses a case similar to the \anon{Tagliavini} collection in his examination of the Benton Fletcher Collection at National Trust Fenton House \cite{McAlpine2014}. When these instruments were donated, Benton Fletcher stipulated that they remain playable and should continue to be maintained for tuition and public performance. A large sampling campaign was conducted, and a custom MIDI interface was designed to fulfil this requirement while preserving the original instruments' integrity. The MIDI controller, comprising two commercially available keyboards mimicking the two-manual harpsichord layout, was used by visitors to trigger the instrument samples recorded with tailored strategies for each. However, user tests identified a limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical keyboard \cite{McAlpine2014}.

@@

The keyboard presented here inherits some aspects from the Tromba Moderna project. The project aims to offer a tool to enhance the fruition of the \anon{Tagliavini} collection while retaining a form of continuity with historical instrument-building traditions. However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms. 
The optical sensing technique for the keyboard is adapted from a similar project on the piano by McPherson \cite{McPherson2013}. 
Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions. This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME -- the methodologies and reasoning for exploring `old' instrument designs at NIME \cite{Masu_NIME_2023}. The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development. Relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.

Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project \anon{\cite{NEMUS}}, aiming to virtually reproduce the sound of historical keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1} investigating embodied relationships between instrument and performer.

@@

The design of the augmented replica keyboard for the \anon{Tagliavini} Collection was guided by three principal constraints stipulated by the museum:

\begin{itemize}
    \item \emph{No visible electronics}. To preserve the visual integrity of the exhibit, all electronic components had to remain concealed except for a pair of headphones and a small display for audio parameter adjustments.
    \item \emph{Robustness and reliability}. The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
    \item \emph{Faithfulness}. The keyboard mechanism ensured fidelity to its authentic operation. The electronics system would not seek to `fix' or `improved' on limitations of the original 16th-century design. 
\end{itemize}

@@

The final design, shown in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key, two-register harpsichord keyboard replica. The keyboard layout was deliberately modelled after typical 16th-century Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}. This effect is particularly relevant in musical instruments, as demonstrated by studies conducted by Fritz \emph{et al.} \cite{Fritz2012, Fritz2014, Fritz2017}. Instead of attempting to mitigate this influence, the design embraces it. This phenomenon, akin to a `musical instrument McGurk effect’, is one of the reasons why the electronics have been concealed from view. The aesthetics of the interface enhance its likelihood of being perceived as an `authentic' musical instrument. Given the extensive research into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020}, the visual component may provide the final persuasive element necessary for acceptance, similar to the way visual perception affects judgments of musical performance \cite{Tsay2013}.

@@

Optical sensors are positioned in front of each jack and their output voltage depends on the amount of light reflected from a sensor surface. In this case, the sensor surface is a grey-scale gradient printed on a vinyl sticker (Figure \ref{fig:jack-tags}) and applied to the side of each jack. Though the current configuration simply sends MIDI note-on and -off messages when passing a threshold, the displacement of all jacks is available continuously.
The jacks for a single key typically have their pluck points offset from each other, a practice known as `staggering'.
A combination of string tension, quill voicing, and jack staggering contribute to the overall `weight' of the action, with each key differing from the low to the high octaves \cite{Veroli2012}. The staggering of the pluck point of the jacks between registers means that there are two distinct tactile anchors along the key dip. The current sensor system enables the data recording of traditional use from which it can be identified where those opportunities for new expression lie \cite{McPherson2013-2}.

@@

\subsection{Materials and Construction}
The keyboard was designed to replicate the tactile and aesthetic sensations of playing an antique Italian harpsichord. Traditional materials were used, including walnut for the wrest plank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard. The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra. The design was inspired by Venetian harpsichords, particularly the \anon{1547 Alessandro Trasuntino} harpsichord at the \anon{San Colombano Museum}. An exception was made for the short octave --- the assigning of common keys in the first octave instead of a chromatic scale --- typically found in Italian harpsichords, which was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.

@@

\begin{figure}
    \centering    
    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
    \caption{3-Key Model Harpsichord Mechanism acquired from \anon{San Colombano, Bologna}, 2023.}
    \label{fig:3key}
\end{figure}

@@

The initial stage of development focused on testing whether sensor data could reliably trigger MIDI playback. Modifying an existing harpsichord for testing was considered but ultimately discarded due to significant internal measurement and layout discrepancies. Instead, a custom 3-key harpsichord mechanism (Figure \ref{fig:3key}) was acquired from the museum and used as a foundation for prototyping. This approach followed a methodology similar to that used in Timmermans \emph{et al.}'s Haptic Key project \cite{Timmermans2020}, since the 3-key model enabled iterative testing of individual components, including sensor placement, signal processing, and mechanical tolerances, before upscaling.

@@

% In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors. 
The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 visitors -- who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
The collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.
The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' found in the Benton Fletcher Collection \cite{McAlpine2014} were absent in this initial feedback.

>>>>

The \anon{Tagliavini} Collection in \anon{Bologna \cite{Tagliavini2007}}, renowned for its historical keyboard instruments, exemplifies this dilemma. Housed at Museo San Colombano - Genus Bononiae, with more than seventy early keyboard instruments, primarily early plucked stringed keyboards of italian origin. The collection stands out as a valuable resource for musicologists, organologists and musicians alike. Preserving the instruments' authenticity was the cornerstone of \anon{Ferdinando Tagliavini}’s vision. This guiding principle led him to collect instruments which could be restored to playing condition after minimal intervention.

However, the delicate mechanisms of these instruments and their historical significance mean they are only played under strict conditions—by experienced historical keyboard performers or by young musicians under strict supervision from the curator. 
To enhance accessibility and engagement, the museum commissioned the replica (Figure \ref{fig:teaser}) of a historical keyboard built in the tradition of the Italian school, which is the subject of this paper.
The keyboard, which museum visitors can use without special permission, presents two choirs of muted strings and can generate MIDI messages. The interface is currently connected to a commercial software sampler, though the ultimate goal is to render accessible the sound of instruments that can not be kept on a playing condition. Museum visitors are invited to play the interface and listen through a pair of headphones. This work outlines the technological aspects of the interface's construction and offers reflections on its role within the \anon{Tagliavini} Collection and potential application within the wider musical instrument museum context.

As part of the initiative to improve the accessibility of museums, the Museo San Colombano - Tagliavini Collection, part of the cultural itinerary of Fondazione Carisbo, has designed a series of projects to improve cognitive and sensory accessibility at different levels, in a project funded by NextGenerationEU. The aim of this project is to exploit the potential of the collection of musical instruments to create an immersive experience and interaction with the heritage on display. 
reinterpretation of the historical musical objects in their tactile-auditory dimension aims to open up the exhibition to blind and partially sighted visitors. The augmented keyboard fulfils many of the project's aims and is part of the initiative to allow visitors of different abilities to engage in a hands-on creative process within the museum.

@@

McAlpine discusses a case similar to the \anon{Tagliavini} collection in his examination of the Benton Fletcher Collection at National Trust Fenton House \cite{McAlpine2014}. When these instruments were donated, Benton Fletcher stipulated that they remain playable and should continue to be maintained for tuition and public performance. A large sampling campaign was conducted, and a custom MIDI interface was designed to fulfil this requirement while preserving the original instruments' integrity. The MIDI controller, comprising two commercially available keyboards mimicking the two-manual harpsichord layout, was used by visitors to trigger the instrument samples recorded with tailored strategies for each. However, user tests identified a limitation: the commercially available weighted keys failed to provide an authentic sense of interacting with a historical plucked keyboard instrument \cite{McAlpine2014}.

@@

The keyboard presented here inherits some aspects from the Tromba Moderna project. The project aims to offer a tool to enhance the fruition of the \anon{Tagliavini} collection while retaining a form of continuity with historical instrument-building traditions. However, the electronics hidden inside the interface are not intended to augment or disrupt the tactile feedback; rather, they serve as the silent and invisible link between the mechanical and the digital realms. 
The optical sensing technique for the keyboard is adapted from a similar project on the piano by McPherson \cite{McPherson2013}. 
Thus, whether the interface proposed here is a \emph{new} musical interface is subject to discussion, particularly regarding the keyboard's tactile response designed to adhere to longstanding harpsichord building traditions. This work complements and follows up on previous reflections within the NIME community, emphasising the \emph{O} in NIME -- the methodologies and reasoning for exploring `old' instrument designs at NIME \cite{Masu_NIME_2023}. 
\textcolor{red}{ALT: This work has incorporated and developed further the ideas of reusing and complimenting the `old' in digital musical instrument design as put forward in `The \emph{O} in NIME' \cite{Masu_NIME_2023}}.
The intended use of the keyboard through meaningful interaction with a museum exhibit is where the novelty of this work lies, rather than solely in its technological development. Relying on previous NIMEs, this project extends the longevity of the results beyond the scope envisaged in the original works, enhancing their sustainability, as discussed by Masu \emph{et al.} \cite{Masu_NIME_2023}.

Besides enhancing the visitor experience, future design iterations will serve as a research tool to explore the unique characteristics of the harpsichord and its impact on performance through the \anon{ERC-funded NEMUS} project \anon{\cite{NEMUS}}, aiming to virtually reproduce the sound of antique keyboard instruments, and upcoming projects such as Rem@ke \cite{remake1} investigating embodied relationships between instrument and performer.

@@

The design of the augmented replica keyboard for the \anon{Museo San Colombano - Genus Bononiae} was guided by three principal constraints stipulated by the museum:

\begin{itemize}
    \item \emph{Faithfulness}. The keyboard mechanism ensured fidelity to its authentic operation. The electronics system would not seek to `fix' or `improved' on limitations of the original design.     
    \item \emph{Robustness and reliability}. The system needed to accommodate frequent use by museum visitors and allow for straightforward maintenance by staff without requiring specialised technical expertise.
    \item \emph{No visible electronics}. To preserve the visual integrity of the exhibit, all electronic components had to remain concealed except for a pair of headphones and a small display for audio parameter adjustments.
    \item \emph{Reduced use of space}. As museums often face space limitations, the instrument needed be compact enough not to compromise space for exhibition of the permanent display. 
\end{itemize}

@@

The final design, shown in Figures \ref{fig:teaser} and \ref{fig:details}, is a 49-key, two-register harpsichord keyboard replica. The keyboard layout was deliberately modelled after early Italian harpsichords, leveraging the human tendency to be influenced by visual elements when making musical judgments \cite{Tsay2013}. This effect is particularly relevant in musical instruments, as demonstrated by studies conducted by Fritz \emph{et al.} \cite{Fritz2012, Fritz2014, Fritz2017}. Instead of attempting to mitigate this influence, the design embraces it. This phenomenon, akin to a `musical instrument McGurk effect’, is one of the reasons why the electronics have been concealed from view. The aesthetics of the interface enhance its likelihood of being perceived as an `authentic' musical instrument. Given the extensive research into modelling and recreating piano action \cite{Cadoz1990, Gillespie1996, Timmermans2020}, the visual component may provide the final persuasive element necessary for acceptance, similar to the way visual perception affects judgments of musical performance \cite{Tsay2013}.

@@

Optical sensors are positioned in front of each jack and their output voltage depends on the amount of light reflected from a sensor surface. In this case, the sensor surface is a grey-scale gradient printed on a vinyl sticker (Figure \ref{fig:jack-tags}) and applied to the side of each jack. Though the current configuration simply sends MIDI note-on and -off messages when passing a threshold, the displacement of all jacks is available continuously.
The jacks for a single key typically have their pluck points offset from each other, a practice known as `staggering'.
A combination of string tension, quill voicing, and jack staggering contribute to the overall resistance of the action, with each key differing from the low to the high octaves \cite{Veroli2012}. The staggering of the pluck point of the jacks between registers means that there are two distinct tactile anchors along the key dip. The current sensor system enables the data recording of traditional use from which it can be identified where those opportunities for new expression lie \cite{McPherson2013-2}.

@@

\subsection{Materials and Construction}
The keyboard was designed to replicate the tactile and aesthetic sensations of playing an historical Italian harpsichord. Traditional materials were used, including walnut for the wrest plank, chestnut for the key levers, boxwood and ebony for key covers, and cypress for the case and soundboard. The 98 jacks were made from beech, fitted with brass springs and natural seagull feather plectra. The design was inspired by Venetian harpsichords, particularly the \anon{1547 Alessandro Trasuntino} harpsichord at the \anon{San Colombano Museum}. An exception was made for the short octave --- the assigning of common keys in the first octave instead of a chromatic scale --- typically found in Italian harpsichords, which was replaced by a standard octave layout to more easily accommodate the interaction with commercial sample libraries.

@@

\begin{figure}
    \centering    
    \includegraphics[width=\linewidth]{src/images/3-key-side.png}
    \caption{
    3-Key Model Harpsichord Mechanism    
    %acquired from \anon{San Colombano, Bologna}, 2023.
    }
    \label{fig:3key}
\end{figure}

@@

The initial stage of development focused on testing whether sensor data could reliably trigger MIDI playback. Modifying an existing harpsichord for testing was considered but ultimately discarded due to significant internal measurement and layout discrepancies. Instead, a custom 3-key harpsichord mechanism (Figure \ref{fig:3key}) was used as a foundation for prototyping. This approach followed a methodology similar to that used in Timmermans \emph{et al.}'s Haptic Key project \cite{Timmermans2020}, since the 3-key model enabled iterative testing of individual components, including sensor placement, signal processing, and mechanical tolerances, before upscaling.

@@

% In an effort to make the exhibition more accessible and authentic, the \anon{San Colombano Museum} commissioned the construction of a mechanical keyboard to trigger a MIDI-controlled sample library, designed for use by museum visitors. 
The exhibition is set to open approximately a month after this work's submission, and only preliminary feedback has been collected from a pool of 20 people -- consisting of staff, surveillance personnel and visitors who were given training on the system -- and expert feedback from \anon{Catalina Vicens} and \anon{Roberto Livi}.
The collection of visitor feedback was carried out as short informal interviews and primarily as a means of identifying any technical problems.
The keyboard shows promise in enhancing the museum experience as comments about a ``sense of disconnect'' found in the Benton Fletcher Collection \cite{McAlpine2014} were absent in this initial feedback.

### 2025-02-04 src/main.tex
Updates from Overleaf
Start catalina comments

%%
%% This is file `sample-nime-paper.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-nime-paper.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% History of NIME Template Updaters:
%% Modified by Florent Berthaut, Charles Martin, Yichen Wang, 21 November 2024
%% Modified by Adnan Marquez-Borbon 30 November 2022
%% Modified by Courtney Reed 28 November 2022
%% Modified by Joe Wright 14 December 2019
%% Modified by Niccolò Granieri 10 October 2018
%% Modified by Angelo Fraietta 23 December 2018
%% Modified by Angelo Fraietta 22 November 2018
%% Modified by Rodrigo Schramm on 22 September 2018
%% Modified by Luke Dahl on 17 October 2-17
%% Modified by Cumhur Erkut on <2016-10-11 Tue>
%% Modified by Edgar Berdahl on 5 November 2014
%% Modified by Baptiste Caramiaux on 25 November 2013
%% Modified by Kyogu Lee on 7 October 2012
%% Modified by Georg Essl on 7 November 2011
%%
%% And here's the template:
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{nimeart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{nimeart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf]{nimeart}
% \documentclass[sigconf,anonymous]{nimeart}
% \documentclass[sigconf,anonymous,review]{nimeart}
\usepackage{graphicx,subcaption,svg}

@@

\author{Roberto Livi}
\affiliation{%
  \institution{Master Harpsichord Maker \\ Museo di San Colombano}
  \city{Pesaro}
  \country{Italy}
  }
\email{robertvs.livi@gmail.com}

\author{Catalina Vicens}
\affiliation{%
  \institution{Museo di San Colombano}
  \city{Bologna}
  \country{Italy}}
  \email{catalina.vicens@genusbononiae.it}

@@

This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument. 
The keyboard was commissioned for exhibition in a musical instrument museum to enhance the visitor experience by providing an interface to digitised versions of instruments within the collection. 
The replica is designed to balance the competing demands of historical authenticity, public accessibility, and preservation. 
It replicates the original instrument's tactile feedback and mechanical resistance by employing historically informed construction techniques. Optical sensors integrated within the mechanism capture the jacks' motion data, enabling MIDI message generation. 
This work situates itself within broader discussions on the role of technology in museums. 
A keyboard interface of this type offers an opportunity to enhance visitor interaction while safeguarding delicate artefacts. 
The paper examines the keyboard's design principles, technical implementation, and implications, emphasising its contribution to public engagement and the long-term preservation of musical heritage.

@@

\section{Ethical Standards}
This work is an empirical study of a keyboard system and
involved no research participants. The authors obtained explicit consent to collect feedback. Seagull feathers, employed as a building material for the jacks, were ethically sourced and picked up along the Pesaro coastline.

>>>>

%%
%% This is file `sample-nime-paper.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-nime-paper.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% History of NIME Template Updaters:
%% Modified by Florent Berthaut, Charles Martin, Yichen Wang, 21 November 2024
%% Modified by Adnan Marquez-Borbon 30 November 2022
%% Modified by Courtney Reed 28 November 2022
%% Modified by Joe Wright 14 December 2019
%% Modified by Niccolò Granieri 10 October 2018
%% Modified by Angelo Fraietta 23 December 2018
%% Modified by Angelo Fraietta 22 November 2018
%% Modified by Rodrigo Schramm on 22 September 2018
%% Modified by Luke Dahl on 17 October 2-17
%% Modified by Cumhur Erkut on <2016-10-11 Tue>
%% Modified by Edgar Berdahl on 5 November 2014
%% Modified by Baptiste Caramiaux on 25 November 2013
%% Modified by Kyogu Lee on 7 October 2012
%% Modified by Georg Essl on 7 November 2011
%%
%% And here's the template:
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{nimeart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{nimeart} or whichever template is required
%% for your publication.
%%
%%
% \documentclass[sigconf]{nimeart}
\documentclass[sigconf,anonymous]{nimeart}
% \documentclass[sigconf,anonymous,review]{nimeart}
\usepackage{graphicx,subcaption,svg}

@@

\author{Roberto Livi}
\affiliation{%
  \institution{Master Harpsichord Maker \\ Museo San Colombano}
  \city{Pesaro}
  \country{Italy}
  }
\email{robertvs.livi@gmail.com}

\author{Catalina Vicens}
\affiliation{%
  \institution{Museo San Colombano}
  \city{Bologna}
  \country{Italy}}
  \email{catalina.vicens@genusbononiae.it}

@@

This paper describes the design and creation of an electronically augmented replica of a harpsichord keyboard with a typical early Italian layout to create a digital musical instrument. 
The keyboard was commissioned for exhibition in a musical instrument museum to enhance the visitor experience by providing an interface to digitised versions of instruments within the collection. 
The replica attempts to balance the competing demands of historical authenticity, public accessibility, and preservation. 
It replicates the original instrument's tactile feedback and mechanical resistance by employing historically informed construction techniques. Optical sensors integrated within the mechanism capture the jacks' motion data, enabling MIDI message generation. 
This work situates itself within broader discussions on the role of technology in museums. 
A keyboard interface of this type offers an opportunity to enhance visitor interaction with cultural heritage while safeguarding delicate artefacts. 
The paper examines the keyboard's design principles, technical implementation, and implications, emphasising its contribution to public engagement and the long-term preservation of musical heritage.

@@

\section{Ethical Standards}
This work is an empirical study of a keyboard system and
involved no research participants. The authors obtained explicit consent to collect feedback. Seagull feathers, employed as a building material for the jacks, were ethically sourced and collected along the Pesaro coastline.
